### 2017-10-30 manuscript.rmd
removed figure size constraints fromm simmap plots

SIMMAP Sex distribution:
```{r sex, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap SEX
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Dioecious", "Monoecious")
sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(sex_sim)
```

@@

SIMMAP Presence of palpons:
```{r palpons, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap PALPONS
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(palpons_sim)
```

@@

SIMMAP Presence of tentilla:
```{r tentilla, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim)
```

@@

SIMMAP Presence of tentilla - *Physalia* corrected:
```{r tentilla2, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim2)
```

@@

SIMMAP Presence of Nectophores:
```{r nectosome, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap NECTOSOME
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(nectosome_sim)
```

@@

SIMMAP Position of Nectosome:
```{r nectosome_position, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#SIMMAP NECTOSOME POSITION
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green")
names(colors) = c("Dorsal", "No nectosome", "Ventral")
necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

SIMMAP Nectophore number and types:
```{r nectophore_number_types, echo=F, message=F,fig.height=7, fig.width=12, warning=F}
#Simmap NECTOPHORE TYPES
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green", "blue")
names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

SIMMAP Habitat:
```{r depth_zones, echo=F, fig.height=7, fig.width=12, message=F, warning=F}
#Simmap Vertical.Range
plotTree(ultratree, lwd = 4)
colors = c("black", "cyan", "blue", "green", "red")
names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

Bathymetrics phylomorphospace
```{r depth_phylomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
par(oma=rep(3,4))
phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
```

Bathymetrics phylomorphospace with time from root as color
```{r depth_phylochronomorpho, echo=F, fig.height=12, fig.width=20, message=F, warning=F}
evotime = fastBM(dpruned_tree,nsim=2)
evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
```

@@

Uncertainty inclusive traitgram for depth range
```{r uncertainty_depth, echo=F, fig.height=14, fig.width=14, message=F, warning=F}
contTrait <- depth_range
A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
trans<-as.character(floor(0:50/2))
trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
# plot
for(i in 0:50){
  p<-i/length(trans)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
}
phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
```

Bathymetrics scattergram array
```{r depth_scattergram, echo=F, fig.height=11, fig.width=15, message=F, warning=F}
fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
```

@@

Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F, fig.height=10, fig.width=18}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=3)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
rownames(f_cdata) = rownames(cdata)
names(f_cdata)[4] = "Float"
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
```

@@

Phylogenetic PCA for bathymetrics
```{r pPCA_depth, echo=F, fig.height=18, fig.width=18, message=F, warning=F}
phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot

>>>>

SIMMAP Sex distribution:
```{r sex, echo=F, message=F, warning=F}
#Simmap SEX
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Dioecious", "Monoecious")
sex_sim %>% plotSimmap(lwd = 4, add = T, colors=colors)
nodelabels(pie=(describe.simmap(sex_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(sex_sim)
```

@@

SIMMAP Presence of palpons:
```{r palpons, echo=F, message=F, warning=F}
#Simmap PALPONS
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
palpons_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(palpons_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(palpons_sim)
```

@@

SIMMAP Presence of tentilla:
```{r tentilla, echo=F, message=F,warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim)
```

@@

SIMMAP Presence of tentilla - *Physalia* corrected:
```{r tentilla2, echo=F, message=F, warning=F}
#Simmap TENTILLA
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
tentilla_sim2 %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(tentilla_sim2, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(tentilla_sim2)
```

@@

SIMMAP Presence of Nectophores:
```{r nectosome, echo=F, message=F, warning=F}
#Simmap NECTOSOME
plotTree(ultratree, lwd = 4)
colors = c("black", "red")
names(colors) = c("Absent", "Present")
nectosome_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(nectosome_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
densityMap(nectosome_sim)
```

@@

SIMMAP Position of Nectosome:
```{r nectosome_position, echo=F, message=F, warning=F}
#SIMMAP NECTOSOME POSITION
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green")
names(colors) = c("Dorsal", "No nectosome", "Ventral")
necto_pos_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_pos_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

SIMMAP Nectophore number and types:
```{r nectophore_number_types, echo=F, message=F, warning=F}
#Simmap NECTOPHORE TYPES
plotTree(ultratree, lwd = 4)
colors = c("black", "red", "green", "blue")
names(colors) = c("Multiple of one type", "None", "Two of one type", "Two of two types")
necto_types_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(necto_types_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

SIMMAP Habitat:
```{r depth_zones, echo=F, message=F, warning=F}
#Simmap Vertical.Range
plotTree(ultratree, lwd = 4)
colors = c("black", "cyan", "blue", "green", "red")
names(colors) = c("Bathypelagic", "Surface", "Mesopelagic", "Epipelagic", "Deep benthic")
vrange_sim %>% plotSimmap(lwd = 4, add = T,colors=colors)
nodelabels(pie=(describe.simmap(vrange_sim, plot=F)$ace) ,piecol=colors,cex=0.35)
add.simmap.legend(colors = colors, x=0.6*par()$usr[1],y=0.3*par()$usr[4],prompt=FALSE)
```

@@

Bathymetrics phylomorphospace
```{r depth_phylomorpho, echo=F, message=F, warning=F}
par(oma=rep(3,4))
phylomorphospace(depthmap$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=depthmap$cols, fsize=1, label = "horizontal")
```

Bathymetrics phylomorphospace with time from root as color
```{r depth_phylochronomorpho, echo=F, message=F, warning=F}
evotime = fastBM(dpruned_tree,nsim=2)
evotime_map <- contMap(dpruned_tree,evotime[,1], plot=F)
evotime_map$cols[]<-rainbow(1001,start=0.7,end=0)
phylomorphospace(evotime_map$tree, cbind(depth_max, depth_min), xlab="Minimum depth (m)", ylab="Maximum depth (m)", colors=evotime_map$cols, fsize=1, label = "horizontal")
#add.color.bar(2,evotime_map$cols,title="Time from the root", lims=c(0,2000),digits=5)
```

@@

Uncertainty inclusive traitgram for depth range
```{r uncertainty_depth, echo=F, message=F, warning=F}
contTrait <- depth_range
A<-fastAnc(dpruned_tree,contTrait,CI=TRUE)
paintree<-paintSubTree(dpruned_tree,node=length(dpruned_tree$tip)+1,"1")
trans<-as.character(floor(0:50/2))
trans[as.numeric(trans)<10]<- paste("0", trans[as.numeric(trans)<10],sep="")
# plot
for(i in 0:50){
  p<-i/length(trans)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,1]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=i>0)
  phenogram(dpruned_tree,c(contTrait,(1-p)*A$CI95[,2]+p*A$ace), colors=setNames(paste("#0000ff",trans[i+1],sep=""),1), add=TRUE)
}
phenogram(dpruned_tree,c(contTrait,A$ace),add=TRUE, colors=setNames("black",1))
```

Bathymetrics scattergram array
```{r depth_scattergram, echo=F, message=F, warning=F}
fancyTree(dpruned_tree,type="scattergram",X=cbind(depth_max, depth_min, depth_range, depth_median),control=list(spin=FALSE))
```

@@

Overview grid of main binary traits
```{r ggtree_bin_heatmap, echo=F, message=F, warning=F}
p <- ggtree(ultratree)
p <- p + geom_tiplab(size=3)
f_cdata = sapply(cdata, as.factor) %>% as.data.frame %>% .[c(2,5:7)]
rownames(f_cdata) = rownames(cdata)
names(f_cdata)[4] = "Float"
gheatmap(p, f_cdata, offset = 0.4, width=0.5, colnames_position = 'top', font.size = 3)
```

@@

Phylogenetic PCA for bathymetrics
```{r pPCA_depth, echo=F, message=F, warning=F}
phyl.pca(dpruned_tree, dpruned_data[,c(14,15,16,18,19)], mode="corr") %>% biplot
